---
title: "Data analysis"
output: html_document
date: "2025-09-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages

```{r}
library("tidyverse")
library("patchwork")
library("RColorBrewer")
```


Import raw data from xlsx files

```{r}
library(readxl)
  Raw_data<-read_excel("Raw_data.xlsx")

``` 


Calculate survival rates. Survival rate is entered as a new vector in existing data frames

```{r}
Data<-Raw_data %>%
   group_by(Rep, Tmt) %>%              # group by tank (or replicate) so each has its own baseline
  mutate(
    Survival = (Ind / first(Ind[Week == 0])) * 100
  ) %>%
  ungroup()
```


Calculate individual weight. Individual weight is entered as a new vector in existing data frames


```{r}
Data<-Data %>%
  mutate(
    W=as.numeric(W),
    Ind_W=W/Ind)
```

Calculate weight gain (WG), percentage weight gain (PWG) and specific growth rate (SGR) on an individual weight basis

```{r}
Data<-Data %>%
   group_by(Rep, Tmt) %>%              
  mutate(
    WG=(Ind_W-first(Ind_W[Week == 0]))
  ) %>%
  ungroup()

Data<-Data %>%
   group_by(Rep, Tmt) %>%              
  mutate(
    PWG=(WG/first(Ind_W[Week == 0]))*100
  ) %>%
  ungroup()

Data<-Data %>%
   group_by(Rep, Tmt) %>%              
  mutate(
    SGR=(log(Ind_W)-log(first(Ind_W[Week == 0])))/14*100
  ) %>%
  ungroup()

```

Calculate weight gain (WG) and percentage weight gain (PWG) on a tank biomass basis

```{r}
Data<-Data %>%
   group_by(Rep, Tmt) %>%              
  mutate(
    WG_TB=(W-first(W[Week == 0]))
  ) %>%
  ungroup()

Data<-Data %>%
   group_by(Rep, Tmt) %>%              
  mutate(
    PWG_TB=(WG_TB/first(W[Week == 0]))*100
  ) %>%
  ungroup()

```

Calculate mean and standard error values for survival and growth indices

```{r}
Mean_survival <- Data %>% #Calculate mean survival rate by treatment and week
  group_by(Tmt, Week) %>%
  summarise(
    Survival = mean(Survival, na.rm = TRUE),
    SE=sd(Survival)/sqrt(3),
    .groups = "drop"
  )
    
  Mean_growth <- Data %>% 
  filter(!(Week %in% c(0,1,3))) %>% #Exclude weeks not containing any growth index data
    group_by(Tmt, Week) %>%
  summarise(
    WG_mean = mean(WG, na.rm = TRUE), #Calculate growth index on an individual weight basis
    WG_SE=sd(WG,na.rm=TRUE)/sqrt(3),
    PWG_mean = mean(PWG, na.rm = TRUE),
    PWG_SE=sd(PWG,na.rm=TRUE)/sqrt(3),
    SGR_mean = mean(SGR, na.rm = TRUE),
    SGR_SE=sd(SGR,na.rm=TRUE)/sqrt(3),
    
    WG_TB_mean = mean(WG_TB, na.rm = TRUE), #Calculate growth index on a tank biomass basis
    WG_TB_SE=sd(WG_TB,na.rm=TRUE)/sqrt(3),
    PWG_TB_mean = mean(PWG_TB, na.rm = TRUE),
    PWG_TB_SE=sd(PWG_TB,na.rm=TRUE)/sqrt(3),
    
    .groups = "drop"
  )
```
---------------------------------------------------------------------------------
Plots

Plotting survival rates over time

```{r}

ggplot(Mean_survival, aes(x=Week, y=Survival, color=Tmt)) +  #Survival rate over time
  geom_point() +
  geom_line() +
  labs(title="Survival rate over time", 
       x="Week", 
       y="Survival Rate (%)") +
  scale_y_continuous(
    breaks=c(0,20,40,60,80,100),
    limits=c(0,100)
    )+
  theme_minimal()+
  theme(plot.title=element_text(size = 14, face = "bold",hjust=0.5))

```


Performance indices over time

```{r}
ggplot(Mean_growth,aes(x=Week,y=WG_mean,color=Tmt))+  #WG over time
  geom_point()+
  geom_line(aes(group=Tmt))+
  theme_minimal()+
  labs(title="Weight gain over time",
       x="Week",
       y="SGR (%/day)")+
  theme(plot.title=element_text(size = 16, face = "bold",hjust=0.5))+
  scale_x_continuous(
    breaks=c(2,4,6),
    limits=c(2,6)
  )

ggplot(Mean_growth,aes(x=Week,y=PWG_mean,color=Tmt))+  #PWG over time
  geom_point()+
  geom_line(aes(group=Tmt))+
  theme_minimal()+
  labs(title="Percent weight gain over time",
       x="Week",
       y="SGR (%/day)")+
  theme(plot.title=element_text(size = 16, face = "bold",hjust=0.5))+
  scale_x_continuous(
    breaks=c(2,4,6),
    limits=c(2,6)
  )


ggplot(Mean_growth,aes(x=Week,y=SGR_mean,color=Tmt))+  #SGR over time
  geom_point()+
  geom_line(aes(group=Tmt))+
  theme_minimal()+
  labs(title="Specific growth rate over time",
       x="Week",
       y="SGR (%/day)")+
  theme(plot.title=element_text(size = 16, face = "bold",hjust=0.5))+
  scale_x_continuous(
    breaks=c(2,4,6),
    limits=c(2,6)
  )

```

Performance indices by treatment

```{r}
Mean_growth %>%                             #PWG at week 2
  filter(Week == 2) %>%
ggplot(aes(x=Tmt,y=PWG_mean,color=Tmt))+
  geom_point()+
  geom_errorbar(aes(ymin=PWG_mean-PWG_SE,
                    ymax=PWG_mean+PWG_SE),width=0.2)+
  labs(title="Percent weight gain at week 2",
       x="Treatment",
       y="Weight gain (%)")+
  theme_minimal()+
  theme(plot.title=element_text(size = 16, face = "bold",hjust=0.5))


Mean_growth %>%                             #PWG at week 4
  filter(Week == 4) %>%
ggplot(aes(x=Tmt,y=PWG_mean,color=Tmt))+
  geom_point()+
  geom_errorbar(aes(ymin=PWG_mean-PWG_SE,
                    ymax=PWG_mean+PWG_SE),width=0.2)+
  labs(title="Percent weight gain at week 4",
       x="Treatment",
       y="Weight gain (%)")+
  theme_minimal()+
  theme(plot.title=element_text(size = 16, face = "bold",hjust=0.5))


Mean_growth %>%                             #SGR at week 2
  filter(Week == 2) %>%
ggplot(aes(x=Tmt,y=SGR_mean,color=Tmt))+
  geom_point()+
  geom_errorbar(aes(ymin=SGR_mean-SGR_SE,
                    ymax=SGR_mean+SGR_SE),width=0.2)+
  labs(title="Specific growth rate at week 2",
       x="Treatment",
       y="SGR (%Growth/day")+
  theme_minimal()+
  theme(plot.title=element_text(size = 16, face = "bold",hjust=0.5))


Mean_growth %>%                             #SGR at week 4
  filter(Week == 4) %>%
ggplot(aes(x=Tmt,y=SGR_mean,color=Tmt))+
  geom_point()+
  geom_errorbar(aes(ymin=SGR_mean-SGR_SE,
                    ymax=SGR_mean+SGR_SE),width=0.2)+
  labs(title="Specific growth rate at week 4",
       x="Treatment",
       y="SGR (%Growth/day)")+
  theme_minimal()+
  theme(plot.title=element_text(size = 16, face = "bold",hjust=0.5))


```

Plotting relationship between # of individuals and individual weight

```{r}
Data %>%
  filter(!(Week %in% c(1,3))) %>%
ggplot(aes(x=Ind,y=Ind_W)) +
  geom_point() +
  geom_smooth() +
  labs(title="Individual weight by # of individuals per tank ",
       x="# of Individuals",
       y="Individual weight (g)") +
  theme_minimal()

```
